Modular Inverse Reinforcement Learning on Human Motion

### 2015-02-11 paper.tex
on rl

\title{Inverse Modular Reinforcement Learning on Human Motion}

@@

\keywords{
enter key words here
}

\acknowledgements{}

@@

Human is able to learn complicated behaviors much faster than machines can do.

We analyze human's behavior of accomplishing various tasks. With the best of our
knowledge, there is no work using inverse reinforcement learning to understand
human motional behavior in the literature.

We conduct this research motivated by two questions.

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{avatar.png}
\caption{The task of collecting targets, avoiding obstacles and following the
path.}
\label{fig:avatar}
\end{figure}

Consider the task illustrated in Figure~\ref{fig:avatar} A). The avatar is asked to
do three sub-tasks simultaneously --- 1) following the path, indicated by the gray
line on the ground, 2) getting targets, the blue cylinders, and 3) avoiding
obstacles, the pink cylinders. This is an experiment design used in the
literature to evaluate modular reinforcement learning \cite{rothkopf2013modular}.

From the reinforcement learning perspective, this task can be decomposed to be
three sub-tasks as described above.  In Figure~\ref{fig:avatar} B), if the agent
knows the distance and angle to an object, he is expected to know the optimal
action to avoid or pursue it.

A human has not done this kind of task before can achieve a good performance
easily, shown in the experiments described later.

\section{Modular Reinforcement Learning}

Value functions is decomposed \cite{koller1999computing}.

 If a human knows the policies
of the sub-tasks, or sub-MDPs, he can accomplish a complicated behavior by
combining the sub-MDPs. That is,
$$Q(s, a) = \sum_i w_i Q_i (s, a)$$
where $Q_i$ is the Q value of the i-th sub-MDP, $w_i$ is the weight of the i-th
sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.

Different weights can yield different performance. Let $w_1, w_2, w_3$ be
weights for the task of target collection, obstacle avoidance, and path
following, respectively. Let $w$ be the vector of $(w_i)_1^n$. An agent with $w
= (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may avoid
the obstacles and follow the path.

@@

\section{Experiments}

We conducted experiments to ask volunteers to accomplish different tasks and
recorded their trajectories. There are four kind of tasks. Task 1, following the
path only, and ignoring other objects. Task 2, following the path, while avoid
the obstacles. Task 3, following the path, while attain targets when possible.
Task 4, following the path, collecting the targets and avoiding obstacles
simutanously.
The human data are collected by the Center for Perceptual Systems in University
of Texas at Austin.

@@

\begin{table}[h]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 2.34 &  2.13\\
\hline
2 & 3.03 &  \bf{0.13}\\
\hline
3 & \bf{10.19} & 2.28\\
\hline
4 & \bf{9.88} &  \bf{0.03}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of human.}
\end{table}

@@

\section{Conclusion and Future Work}

Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. However, we adopt the weighted sum approach
because this is more reasonble for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscination between these two
modules.

@@

>>>>

\title{Modular Inverse Reinforcement Learning on Human Motion}

@@

\keywords{
Reinforcement learning, human cognition, inverse learning
}

@@

Human is able to learn to accomplish complicated tasks much faster than machines
can do. However, most of the tasks can be decomposed into subtasks. A human may
already have the capacity to accomplish the subtasks. He simply transfers the
knowledge of the subtasks and integrate them for the objective task. For example,
when a human learns to drive, he has much prior domain knowledge to help him
with this objective task. He integrates his skills of controlling the velocity,
avoiding other vehicles, navigating, responding to traffic lights, and so on.

Similar ideas are also adopted in the learning literature. Reinforcement learning
suffers from the curse of dimensionality. It is inefficient for a learning agent
to learn a complicated task from scratch. So learning algorithms including
hierarchical reinforcement learning \cite{dietterich2000hierarchical}, modular
reinforcement learning \cite{sprague2003multiple} are proposed. They also
decompose the objective task into subtasks. The learning algorithms first learn
the subtasks independently, then learn how to combine these subtasks.

A natural question to ask is how human integrates skills for subtasks, and
whether we can use the same method for integrating subtasks in reinforcement
learning problems. In this paper, we analyze human's behavior of
accomplishing a task composite of various subtasks. We collected human motion
data when human accomplish a task, and try to interpret the behavior using
inverse reinforcement learning. With the best of our knowledge, this approach is
noval in the literature.

This paper is organized as follows. Section~\ref{sec:domain} introduces the
domain of the composite task that we collected human data. Section~\ref{sec:rl}
describes the main algorithm, modular inverse reinforcement learning. We report
our experiment results in Section~\ref{sec:exp}, and conclude in
Section~\ref{sec:conclude}.

\section{Multi-objective Sidewalk Domain}
\label{sec:domain}

\begin{figure}[h!]
\centering
\includegraphics[width=0.6\textwidth]{avatar.png}
\caption{The task of collecting targets, avoiding obstacles and following the
path. \cite{rothkopf2013modular}}
\label{fig:avatar}
\end{figure}

Figure~\ref{fig:avatar} shows the domain that we use in this paper. The pink
cylinders represent obstacles, that the player would be punished if running
into. The blue cylinders represent targets, that the player would be rewarded if
collected. There is also a grey path on the ground that the player can follow.
So, naturally, this domain has three subtasks, 1) following the path, 2)
collecting targets, 3) avoiding obstacles. This is an experiment used in the
literature to evaluate modular reinforcement learning
\cite{rothkopf2013modular}.

We will evaluate four different tasks. Task 1, following the path only, and
ignoring other objects. Task 2, following the path, while avoid the obstacles.
Task 3, following the path, while attain targets when possible.  Task 4,
following the path, collecting the targets and avoiding obstacles
simultaneously.
We conducted experiments to ask volunteers to accomplish different tasks and
recorded their trajectories. The human data are collected by the Center for
Perceptual Systems in University of Texas at Austin.

If the player observes the distance and angle to an object, as shown in
Figure~\ref{fig:avatar} B), he is expected to know the optimal action to avoid
or pursue it. This decision is also Markov. Therefore, the question would be, if
we have trained modules for these three subtasks, could they be integrated to
match human's behavior?

\section{Modular Inverse Reinforcement Learning}
\label{sec:rl}

We need some basic concepts in reinforcement learning to proceed. For a state, action
pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
state $s$. The {\em policy} of a state is the action with the maximum Q
value\cite{rl}. So how to we determine the global policy from the modules, or
subtasks? In the literature, work has been done to integrate the Q function
\cite{koller1999computing} or compromise on policies directly
\cite{thomas2012motor}.

Here, we assume that the global Q function is a weighted sum of all $Q_i$, where
$Q_i$ is the Q function for i-th module.
$$Q(s, a) = \sum_i w_i Q_i (s_i, a)$$
where $w_i$ is the weight of the i-th sub-MDP. $w_i \geq 0, \sum_i w_i = 1$.
$s_i$ denotes the decomposition of $s$ in the i-th module.

Different weights can yield different performance. Let $w_1, w_2, w_3$ be
weights for the task of target collection, obstacle avoidance, and path
following, respectively. Let $w$ be the vector of $(w_1, w_2, w_3)$. An agent
with $w = (1, 0, 0)$ only collect targets, and one with $w = (0, 0.5, 0.5)$ may
avoid the obstacles and follow the path.

@@

\section{Experiments}
\label{sec:exp}

@@

\begin{table}[h]
\centering
\begin{tabular}{| l| l| l |}
\hline
Average by Task & Num Targs Hit & Num Obst Hit\\
\hline
1 & 2.34 &  2.13\\
\hline
2 & 3.03 &  \bf{0.13}\\
\hline
3 & \bf{10.19} & 2.28\\
\hline
4 & \bf{9.88} &  \bf{0.03}\\
\hline
\end{tabular}
\caption{Number of targets hit and number of obstacles hit of the humans.}
\end{table}

@@

\section{Conclusion and Future Work}
\label{sec:conclude}

Weighted sum of Q function is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. This is also called skilled in the literature
\cite{konidaris2009skill}. However, we adopt the weighted sum approach
because this is more reasonable for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscillation between these two
modules.

@@

Weights may be dynamic. However, with such assumption, we need to learn a
mapping from state to weights. In this case, the curse of dimensionality still
exists, and inverse learning would be difficult.
